Parsing Postfix log files

### 2008-04-11 doc/logparser.tex
Improved the "aborted delivery attempts" explanation. The heuristics used for detecting them still need work.

XXX FIGURE OUT THE CAUSE OF THE ABORTED DELIVERY ATTEMPTS IN SECTION
5.8.2\@.

@@

Some mail clients appear\footnote{Due to privacy concerns no attempt has
been made to identify either the users or the software which exhibits this
behaviour.} to send the following sequence of commands during
the \SMTP{} session:

\begin{verbatim}
    EHLO client.hostname
    MAIL FROM: <sender@example.com>
    RCPT TO: <recipient@example.com>
    RSET
    MAIL FROM: <sender@example.com>
    RCPT TO: <recipient@example.com>
    DATA
\end{verbatim}

The client aborts the first delivery attempt after the first recipient is
accepted, then makes a second delivery attempt which it continues with
until the delivery is complete.

Once again Postfix does not log a message making the client's behaviour
clear, so once again heuristics are required to identify when this
behaviour occurs.  In this case a list of all mails accepted during a
connection is saved in the connection state, and the accepted mails are
examined when the disconnection action is executed.  Each mail is checked
for the following: is the mail missing its \daemon{cleanup} log message?
Every mail which passes through Postfix will have a \daemon{cleanup} line;
lack of a \daemon{cleanup} line is a sure sign the mail didn't make it too
far.

>>>>

@@

Some mail clients behave unexpectedly during the \SMTP{} dialogue: the
client aborts the first delivery attempt after the first recipient is
accepted, then makes a second delivery attempt for the same recipient which
it continues with until delivery is complete.  An example dialogue
exhibiting this behaviour is shown presented below (lines starting with a
three digit number are sent by the server, the other lines are sent by the
client):

\begin{verbatim}
220 smtp.example.com ESMTP
EHLO client.example.com
250-smtp.example.com
250-PIPELINING
250-SIZE 15240000
250-ENHANCEDSTATUSCODES
250-8BITMIME
250 DSN
MAIL FROM: <sender@example.com>
250 2.1.0 Ok
RCPT TO: <recipient@example.net>
250 2.1.5 Ok
RSET
250 2.0.0 Ok
RSET
250 2.0.0 Ok
MAIL FROM: <sender@example.com>
250 2.1.0 Ok
RCPT TO: <recipient@example.net>
250 2.1.5 Ok
DATA
354 End data with <CR><LF>.<CR><LF>
The mail transfer is not shown.
250 2.0.0 Ok: queued as 880223FA69
QUIT
221 2.0.0 Bye
\end{verbatim}

Once again Postfix does not log a message making the client's behaviour
clear, so once again heuristics are required to identify when this
behaviour occurs.  In this case a list of all mails accepted during a
connection is saved in the connection state, and the accepted mails are
examined when the disconnection action is executed.  Each mail is checked
for the following: is the mail missing its \daemon{cleanup} log message,
and are there exactly two \daemon{smtpd} log lines?  Every mail which
passes through Postfix will have a \daemon{cleanup} line; lack of a
\daemon{cleanup} line is a sure sign the mail didn't make it too far.  The
two \daemon{smtpd} log lines will be a connection log line and a mail
acceptance log line (Postfix logs acceptance as soon as the first recipient
has been accepted); XXX IMPROVE THIS ONCE I HAVE BETTER HEURISTICS\@.
